Fish do not need Visas: the transboundary nature of world fished species

### 2019-11-28 Manuscript_Draft.Rmd
Addressed Coletts changes and comments until Fig 1.

---
title: "Fish do not need Visas: the transboundary nature of world fished species"
author: "Authors: Juliano Palacios-Abrantes^1^, Gabriel Reygondeau^1,2^, William W.L. Cheung^1^"
date: "Affiliations: ^1^ Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada, ^2^ Yale"
subtitle: "DRAFT DO NOT SHARE"
# subtitle: "Fish do not need Visas (16/40 character maximum)"
output:
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
# bibliography: ./References/FishForVisa_Ref.bib
# csl: ./References/science.csl
---

@@

Species are not randomly distributed in the world. The area in space and time where a stable population lives and reproduces is determined by the species' fitness response to a series of biotic and abiotic factors  and their influence in the evolutionary process [@Hunch]. Species distributions are key elements in marine resources management as human-made spatial boundaries are a major part of the management system [@Song:2017iua]. However, incorporating the concept of species distribution has not allays been the case in fisheries management, nor policy [@Alexa]. The delimitation of Economic Exclusive Zones (EEZs) in the early 80's virtually spitted the distribution of many species, creating shared species between nations, and adding another level of complexity to fisheries management. However, 40 years after their establishment, the question of how many shared species exists in the world based on their distribution is still unanswered. Answer this question is a key step in creating suitable management plans that allows a sustainable use of marine resources.

From 1973 to 1982 the members of the United Nations held a series of meetings to discuss regulations regarding the high seas, a region of international common property, which, at that time, represented waters after 12 miles from shore [@McRae and Munro, 1989]. The establishment of the UN Convention on the Law of the Sea (UNCLOS) allowed coastal states to claim jurisdiction over 200 nautical miles of their coasts [@UN:1982]. Despite intended to improve fisheries management by granting property rights over shared resources, this arbitrary delimitation of management areas had no input from bio-geography; ignoring species distributions created what we know today as shared stocks. The Food and Agricultural Organization (FAO) recognizes three types of shared stocks: (*i*) transboundary, those stocks that are shared by 2 or more neighboring coastal nations; (*ii*) straddling, those stocks that are shared by two or more nations and also the high seas; and (*iii*) highly migratory stocks; species that are found in the EEZs of coastal nations that are not necessarily neighbors, and the high seas. The origination of shared stocks required the establishment of new fisheries management methods. Managers and scientists adopted a game theory approach where collaboration would most likely result in the best overall outcome for nations sharing a common resource [@Bailey:2010ga]. Lack of collaboration in shared stocks threatens the stock's sustainability [@REF], reduces the profitability potential of the fishery [@REF], and creates conflict between coastal nations [@REF-Jessica]. This is also true when considering the emminent impacts of climate change on the distribution of marine species around the globe [@SumailaLamCheung]

There is a gap in assessing transboundary species in the world based on their spatial distribution. It was estimated in 1997 that there could be up to 1,500 transboundary fish species in the world, however, such estimation lacked a proper assessment due to limited information at the time [@Caddy:1997ue]. More recently, it was estimated there were 344 shared species, representing a capture of 34.2 × 10^6^ t and a global landed value of USD 30.7 × 10^9^ [@Teh:2015gd]. While this represents a good baseline to start, the analysis was based on a literature review of shared species and did not include a mechanistic way of determine the transboundary nature of the species [@Teh:2015gd]. And so, despite recent research highlighting the interconnection of marine species [@Nandini], transboundary species are yet to be identified at a global scale. This is of uppermost need as the fisheries management approach differs substantially between EEZ-discreet and shared species [@Mills]. Moreover, international conflicts over shared resources are expected as climate change shifts the distribution of marine species and new shared fisheries emerge [@Pinsky:2018cb].

@@

In here, we overlap the known distribution of **968** marine fish species with 248 Economic Exclusive Zones of 175 countries to map the number of transboundary species and and their contribution to global catch and landed value. We adopted the UN definition of transboundary stocks (shared by neighboring EEZs) but carried the analysis at the species level, rather than stocks (populations within a species), due to lack of stock-specific spatial and ecological information on all of the fished species [@TEH].

We combined multiple species distribution data-sources and developed a series of criteria to determine if a species was transboundary. Current species distributions were obtained from an observational database of world marine species [@REFNEreus] and two structurally different species distribution models [@REFGabs; @Zeler]. The first criteria to determine if a species is transboundary was related to the agreement of the data-sources regarding the presence of a species in a single grid-cell. The second criteria was established to confirm the presence of a species by using a specially detailed catch data [@Zeler]. Finally, the third criteria accounted for the spatial distribution of the species between the neighboring EEZs. In addition to present the number of transboundary stocks, we estimated their contribution in terms of fishing revenue and landings and categorize them depending on their catch trend [@Pauly 2012]. Our study provides a first empirical approximation to understand the true transboundary nature of the world fisheries, a world that is deeply connected by larval dynamics [@nandini], fishing fleets [@Kurosawa_Juanito], and shared management [@RFMOS].

@@

As Caddy (1997) pointed out, virtually all coastal nations share at least one transboundary species with their neighbors with some sharing over 100 species with surrounding EEZs (Fig. 1). With the current state of knowledge of fished species and their distributions, we estimate a total of 633 transboundary species in the world, more than double of what had been previously estimated [@Teh:2015gd]. These fisheries are important in terms of both landings and revenue. In the last ten years, the global fleet has landed an average of XXX tones, generating USD 79.7 million in fishing revenue from shared stocks fished within EEZs. This represents **X**% of the global catch and 49.8% of total revenue generated by those fisheries in all EEZs of the world. All these values are higher than previously estimated [@Teh]. The importance of these species exceeds in EEZs like the United States where 116 species are responsible for up to 12.20 millions USD and Peru where 28 species are responsible for USD 5.2 million  (Fig 1). Transboundary fisheries are vast and economically important in regions with historic fisheries-related conflicts. The Asian region has seen 43.0% of all international fisheries conflicts since 2000; Europe, where 11.5% of fishing conflicts occur since 1974s [@Spijkers:2019cz], has been involved in a conflict between Ireland and the European Union over mackerel since 2007 [@Spijkers:2017ij], and finally, the Mediterranean has seen conflicts since the 1980's [@Kliot:1989kv]. The identification of current transboundary species is a first step towards establishing joint management plan of marine resources [@Link:2010ei], key to not only sustain fisheries but also avoid international conflict.

@@

![**Number of transboundary species by EEZ and their production by country**](./Figures/Fig1.png)

The two regions of the world with most shared species (North East and Central Atlantic) present quite different management scenarios. For the North East Atlantic, fisheries are managed by the European Union and quotas allocated by species and regulatory area, with some of them (e.g. Iceland and Norway) potentially having good fisheries management [@Melnychuk:2017gh]. In the other hand, fisheries along the west coast of Africa need to be improved, with countries like Nigeria where stock status, management, and enforcement is lacking [@Melnychuk:2017gh]. Ironically, this region has long been exploited by foreign European fleets with questionable fishing rights systems [@Mills:2004dj] and even some illegal activities [@Belhabib:2015bl]. Thus, just having joint management plans for shared stocks between neighbors will not be enough to achieve the sustainable goals of SDGs [@]. Countries need to work together towards a more equitable system of global fishing rights, combat illegal fishing both within their EEZs and outside them, and address global threats such as climate change and overfishing.

@@

Management of transboundary fisheries needs to be looking forward. As species shift their known distributions to cope with a changing ocean, new transboundary species will arose, challenging the management of shared stocks worldwide [@Pinsky]. Conflicts in regions with high abundance of transboundary species will likely be exacerbated by climate change, as already seen in the dispute of Ireland and the European Union over Mackerel (*species*) [@Spijkers]. Understanding how climate change will affect the biogeography of marine species is a key element for proper spatial-based fisheries management, not only at the EEZ level, but also other levels of spatial tools [@Alexa, @Song:2017iua].

@@

Transboundary species are globally distributed with some regions presenting a high concentration of shared species. Their global contribution to fishing revenue and capture is compelling, with transboundary species playing an important role in many fishing nations. Some of these regions like southern Asia and western Africa, are historically marked by fisheries conflicts becoming hotspots for achieving jointly management of shared species. Climate change is likely to have important effects in the management of these species as they shift their distributions following their ideal environment. Identifying transboundary species within neighboring EEZs is the first step towards a joint management that can result in sustainable fisheries management.

@@

In here we present a novel method to estimate whether or not a species is transboundary. Our method is based on a series of criteria to be met and the overlap of species distributions and the spatial boundary of the world's EEZs. For the current analysis we work at the species level and adopt the Food and Agricultural Organization (FAO) definition of transboundary species as those that occur within the Exclusive Economic Zone (EEZ) of neighboring countries [@Munro:2002uf; @FAO]. All data was scaled at a 0.5 degree latitude x 0.5 degree longitude grid cells (*n* = ~180,000 grid cells).

@@

To determine the pool of transboundary marine species exploited by fisheries within each of the world's EEZs we first extract all the commercial marine fish or invertebrate species from the *Sea Around Us* database (http://www.seaaroundus.org and determined their current distributions. For this, we used four species-distributions data sources based on (*i*) observational data, (*ii*) an Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (**Table x**). Each source represent a different method of estimating the distribution of a given species and thus, provides a more robust result. Only those commercial fished species that had data from all four sources were included in the analysis, and therefore the final dataset comprised XXX species, 100% of the reported marine species in the SAU database and XX% of the reported taxa of the FAO (**See S1_Data**).

*i* The occurrence data was collected from five publicly available repositories; Fishbase (http://fishbase.org), the Global Biodiversity Information Facility (GBIF; https://www.gbif.org/), the Ocean Biogeographic Information System (OBIS; https://obis.org/),  the Intergovernmental Oceanographic Comission (IOC; http://ioc-unesco.org), and the International Union for Conservation of Nature (IUCN; https://www.iucn.org/technical-documents/spatial-data) [@Reygondeau:2019uh].

@@

*ii* The ENM-Nereus consists in a multimodel approach based on a Bioclim and a Bososted Regression Tree model [@Thuiller:2009gp], a Maxtent model [@Phillips:2006ff], and a Non-Parametric Probabilistic Ecological Niche Model [@Beaugrand:2011fd]. Environmental variables modeled were sea surface temperature, surface pH, surface oxygen concentration, and vertically integrated (0–100 m) net primary production (NPP) [@Asch:2018ca]. The ENM-Nereus employed all of the observational data-sources previously mention. Global environmental conditions were obtained and results averaged from three Earth System Models developed by the Geophysical Fluid Dynamics Laboratory's (GFDL- https://www.gfdl.noaa.gov/earth-system-model/), the Institute Pierre Simon Laplace (IPSL- www.icmc.ipsl.fr/), and the Max Planck Institute for Meteorology (MPI- www.mpimet.mpg.de/en/science/models/). See [@Asch:2018ca; @Reygondeau:2019uh] for model's details.

*iii* The SDM-SAU model follows a five-steps process based on species-specific life history information, rather than environmental variables [@Close:2006ux; @Palomares:wg]. For each commercial fish species, the model first uses the FAO major fishing areas and countries EEZs to determine a broad distribution. It then uses life history information to delimit its range within the FAO fishing area (e.g. thermal preference, depth limits). The range is delimited even further by expert-review polygons and compared with that of AquaMaps [@Kaschner:2016tl], OBIS and GBIF occurrence data. The model then determines a species habitat preference based on the assumptions that the relative abundance of a species is determine by the number of habitats in a grid cell and the distance of the species to each habitat, and the importance of the habitat to the species size. Finally, the species equatorial submergence (e.g. the latitudinal region where a species is not seen in between poles) is estimated for each species. See [@Close:2006ux; @Palomares:wg] for model's details.

*iv* The previous models combine observational data with a series of biotic and abiotic information to determine the probability of a species to be in a given space and time. However, this does not mean that the species in question will actually be there. Although the models used have their own methods of double-check species distributions (e.g. ENM-Nereus uses four different species distribution algorithms and SPD-SAU verifies with other models), we used a fourth data-source to corroborate the models' outputs. The *Sea Around Us* estimates catch that was not originally reported in the FAO database for each country based on all publicly available information sources. Fishing catch are also spatially allocated following a country-by-country catch reconstruction and spatial layers according to each species [@Zeller2016]. Thus, we used the SAU catch reconstruction from 1970 to 2000 as the fourth dataset to estimate transboundary species and was also used to estimate their catch contribution within EEZs (average 1970-2000).

@@

We developed a four-criteria method for determining whether or not a species was transboundary. Only species that meet all criteria were considered as transboundary species and the analysis was done only within the EEZs of coastal states. In some cases, criteria also work as indexes to measure the uncertainty in the analysis. All of the analysis was done in the statistical software *R version 3.5.2 (2018-12-20)* and code is available at https://github.com/jepa/FishForVisa.

@@

We estimated the economic contribution of transboundary species for each country using global ex-vessel price [@Tai2017]. The data includes ex-vessel price from multiple sources and a structured interpolation method (e.g. similar countries, species) to fill in data gaps [@Sumaila:2015uc]. The contribution of transboundary species was done by aggregating the catch of each species that happened within the EEZ of the countries sharing that species. For this study, we did not include catch from areas beyong national jurisdiction. Finally, we use the catch data to determine the exploitaiton category of each sepcies within a EEZ. Although this method has previouslly used to estimate stock status [@Grainger:1996tk; @Kleisner:2011wn], the categories presented here are intended to be seen as catch trends, and not status of each species. We only assessed species within each EEZ that had at least ten years of data between the first and last reported landings with at least five consecutive data years. Three categories were drown from the method depending on the year's catch, the maximum catch over time and the minimum catch after the catch peak (**Table XX**). Finally, we reported the predominant category over the last ten years of catch data (2004-2014).

>>>>

---
title: "Fish do not need Visas: the transboundary nature of world fished species"
author: "Authors: Juliano Palacios-Abrantes^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada, ^2^Yale University, New Haven, Connecticut, USA, ^3^Stockholm Resilience Centre, Stockholm University, Stockholm, Sweden"
subtitle: "DRAFT DO NOT SHARE"
# subtitle: "Fish do not need Visas (16/40 character maximum)"
output:
  pdf_document:
    fig_caption: yes
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
geometry: margin=1in
editor_options:
  chunk_output_type: inline
# bibliography: ./References/FishForVisa_Ref.bib
# csl: ./References/science.csl
---

@@

Species distribution around the world is not a random process. The area in space and time that a species population occupies is a response to the availability of resources, a series of biotic and abiotic factors, as well as their influence on the evolutionary processes [@Hunch]. Marine resource management is predicated on knowledge of species distributions as human-made spatial boundaries are inherent to existing management systems (e.g., marine protected areas) [@Song:2017iua]. However, incorporating the concept of species distribution into fisheries management, or policy, is a relatively recent construct [@Alexa]. The delineation of Economic Exclusive Zones (EEZs) in the early 80s, for instance, virtually established boundaries across the distribution of many species, creating shared species between nations, and adding another level of complexity to fisheries management. Forty years after their formal adoption, the question of how many shared species exist in the world based on their distribution is still unanswered. Fishing quotas set by different nations for a shared fishery resource have led to conflict [@Spijkers:2017ij]. Climate change, which is shifting species distributions, is likely to exacerbate this type of conflict [@Pinsky:2018cb] and presents further challenges for fisheries management (@Mills). Thus, answering the questions of how many shared stocks exist in the world, and where these are located, is critical to developing effective and resilient management plans for the continued and future sustainable use and conservation of marine resources.

From 1973 to 1982, members of the United Nations held a series of meetings to discuss regulations regarding the high seas, a region of international common property, at that time, consisting of waters from 12 miles from shore [@McRaeandMunro1989]. The establishment of the UN Convention on the Law of the Sea (UNCLOS) allowed coastal states to claim jurisdiction over the exploration and exploitation of marine resources over 200 nautical miles off their coasts [@UN:1982]. While intended to improve fisheries management by granting property rights over shared resources, this arbitrary delimitation of management areas was not informed by bio-geography, essentially ignoring species distributions and creating what we know today as shared fish stocks [@Gulland, @Munro2002]. Fish stocks, are comonly refered in fisheries management as a particular population that is more or less isolated from other stocks of the same species and hence self-sustaining [@FAO:1999tk]. The Food and Agricultural Organization (FAO) recognizes three types of shared stocks: (*i*) transboundary stocks, shared by 2 or more neighboring coastal nations; (*ii*) straddling stocks, occurring in two or more adjacent national jurisdictions and the high seas; and (*iii*) highly migratory stocks, found in the EEZs of coastal nations that are not necessarily adjoining, and the high seas. The inception of shared stocks called for the establishment of new fisheries management methods. Managers and scientists adopted a game theory approach – or means to analyze strategic interactions among decision-makers – which demonstrates that collaboration is most likely to result in the best overall outcome for nations sharing a common resource [@Bailey:2010ga,@Eide:2013dh]. Lack of collaboration in shared stocks may threaten stock sustainability [@Clark:1980js; @Nguyen:2018dk], reduce the profitability potential of the fishery [@Merino:2007jz], and lead to conflict between coastal nations [@Spijkers:2017ij], although cooperation and non-cooperation can be more complex [@Jensen:2015cf]. Moreover, under climate change, species distributions will shift [@Cheung2016], resulting in new shared stocks [@Pinsky:2018cb] and highlighting the key role cooperative fisheries management can play in maintaining stocks and profitability [@Nguyen:2018dk] as well as hopefully reducing fisheries conflict [@Spijkers:2019cz].

There is a gap in assessing the nature and number of transboundary species in the world based on their spatial distribution. Caddy [@Caddy:1997ue] in 1997 estimated that there could be up to 1,500 transboundary fish species in the world. However, such estimation lacked a proper assessment due to limited information on the distribution of species at the time [@Caddy:1997ue]. More recently, Teh et al. [@Teh:2015gd] estimated that 206 species can be considered shared, accounting for a total catch of 34.2 million t and a global landed value of USD 30.7 trillion. While this represents a good starting point, the analysis was based on a literature review of shared species and did not include a mechanistic way of determining the transboundary nature of the species [@Teh:2015gd]. Thus, despite recent research highlighting the interconnection of marine species [@Ramesh:2019va], the total number of transboundary species is yet to be identified at a global scale. Yet, such an understanding is critical to effectively managing fisheries, as needs differ substantially between discreet and shared species [@Mills] and climate change will yield new challenges associated with new shared fisheries [@Pinsky:2018cb].

@@

In this study, we overlay the known distribution of **968** commercially valuable marine fish species and the Exclusive Economic Zones of 175 coastal countries to determine the total number of transboundary species and their contribution to global catch and landed value. We adopt the UN definition of transboundary stocks (shared by neighboring EEZs), but carry out the analysis at the species level, rather than stocks (populations within a species), due to lack of stock-specific spatial and ecological information on all of the fished species [@Teh:2015gd].

We combined multiple species distribution data-sources and developed a series of three criteria to determine if a species is transboundary. Current species distributions were obtained from the largest observational database of the world’s marine species [@REFNEreus] and two structurally distinct species distribution models [@REFGabs; @Zeler]. The first criteria to determine the transboundary nature of a species considered the number of data-sources agreeing that a species was present in a single grid-cell. Second, we used spatially explicit catch data [@Zeller] to confirm the occurrence of a species in a given grid cell. The species was droped if it was present in any specific grid cell, but no catch was record was found. The third criteria considered the spatial distribution of species between neighboring EEZs. In addition to determining the number of transboundary stocks, we also estimated the contribution of each to fishing revenues and landings as well as categorized them according to catch trends [@Kleisner:2013gh]. Our study provides a first empirical approximation of the number of shared commercial species globally, significantly contributing to our understanding of the true transboundary nature of world fisheries and hopefully helping inform the effective management of these resources.

@@

We estimate a total of 633 transboundary species in the world, more than double the number previously estimated [@Teh:2015gd] (Fig. 1). We also find that fisheries on these species contribute substantially to both landings and revenue. Between 2010 and 2016, fleets targeting shared stocks within EEZs landed an average of **X** tonnes, representing **X%** of the global catch (within EEZs), and generating USD 79.7 million in fishing revenue, equivalent to 49.8% of total revenue. These values are higher than previously estimated [@Teh]. The importance of these species exceeds in EEZs like the United States where 116 species are responsible for up to 12.20 million USD and Peru where 28 species are responsible for USD 5.2 million (Fig 1).

We found transboundary fisheries to be of substantial economic importance in regions with historic fisheries-related conflicts. The Asian region with a total number of **xxx** transboundary species contributing to **xx**% to the region's total landings and **xx**% of total revenue has experienced 43.0% of all documented international fisheries conflicts since 2000 [@Spijkers:2019cz]. The European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for mackerel since 2007 [@Spijkers:2017ij], a species that is transboundary and contributes **xx** to landings and **xx** in revenue to the countries involved in the dispute. The Mediterranean, a region that has experienced conflicts since the 1980s [@Kliot:1989kv], is home to a large number of transboundary stocks that contribute **xx**% to total landings and **xx**% of revenue of countries along the Sea's rim. The identification of current transboundary species is a first step towards establishing joint management plans for marine resources [@Link:2010ei], key not only to sustain fisheries over the long term, but also to avoid international conflict.

@@

![**Number of transboundary species by EEZ and their revenue by country**](./Figures/Fig1.png)

The two regions of the world that share the most species (the North East and Central Atlantic) have developed quite different approaches to managing shared stocks. In the North East Atlantic, fisheries are managed by the European Union and quotas allocated by species and regulatory area, with some of them (e.g., Iceland and Norway) characterized as having good fisheries management [@Melnychuk:2017gh]. On the other hand, fisheries along the west coast of Africa could benefit from improvements [@Melnychuk:2017gh]. Ironically, this region has long been exploited by foreign fleets, including European, sometimes illegally [@Belhabib:2015bl], with questionable fishing rights systems [@Mills:2004dj], and unfair fisheries agreements [@Kaczynski & Fluharty 2002]. Thus, having joint management plans for shared stocks between neighbors will not be enough to achieve sustainable goal 14, Life Below Water [@SDG]. Countries need to work together towards a more equitable system of global fishing rights.

@@

Management of transboundary fisheries needs to be looking forward. As species shift their known distributions to cope with a changing ocean, new transboundary species will arise, challenging the management of shared stocks worldwide [@Pinsky]. Conflicts in regions with a high abundance in transboundary species will likely be exacerbated by climate change, as seen in the ongoing dispute between the European Union, Norway, Iceland and the Faroe Islands over mackerel (**species**) [@Spijkers:2017ij]. Understanding how climate change will affect the biogeography of marine species, and notably the number of countries' EEZ it spans is a key element to the design of effective long-term spatial-based fisheries management [@Alexa, @Song:2017iua].

@@

The global contribution of transboundary species to fishing revenue and capture is compelling, with transboundary species playing an important role in many fishing nations. Transboundary species have played and continue to play an important role in fisheries conflicts, with climate change likely to exacerbate this trend given expected shifts in species distribution and the emergence of 'new' shared species. Identifying existing transboundary species is the first step towards joint management frameworks that are precautionary, strive for sustainability and can be flexible to accommodate future changes.

@@

We present a novel method to determine whether or not a species can be considered transboundary. The method is based on a series of criteria to be met and the overlay of species distributions and the spatial boundaries of the world’s EEZs. For the current analysis we work at the species level and adopt the Food and consider transboundary species as those that occur within the Exclusive Economic Zone (EEZ) of neighboring countries [@Munro:2002uf; @FAO]. All data was scaled to 0.5 degree latitude x 0.5 degree longitude grid cells (*n* = ~180,000 grid cells).

@@

To determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs we first extracted all the commercial marine fish or invertebrate species from the *Sea Around Us* database (http://www.seaaroundus.org and determined their current distributions. For this, we used four data sources of species-distributions: (*i*) observational data, (*ii*) an Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (**Table x**). Each source represent a different method of estimating the distribution of a given species and thus, provides a more robust result. Only commercial fished species with data from all four sources were included in the analysis. The final dataset comprised **XXX** species, **XX**% of the reported marine species in the SAU database and **XX**% of the reported taxa of the FAO (**See S1_Data**).

### *Occurence data*

The occurrence data was collected from five publicly available repositories; Fishbase (http://fishbase.org), the Global Biodiversity Information Facility (GBIF; https://www.gbif.org/), the Ocean Biogeographic Information System (OBIS; https://obis.org/),  the Intergovernmental Oceanographic Comission (IOC; http://ioc-unesco.org), and the International Union for Conservation of Nature (IUCN; https://www.iucn.org/technical-documents/spatial-data) [@Reygondeau:2019uh].

### *Distribution models*

@@

The ENM-Nereus consists in a multimodel approach based on a Bioclim and a Bososted Regression Tree model [@Thuiller:2009gp], a Maxent model [@Phillips:2006ff], and a Non-Parametric Probabilistic Ecological Niche Model [@Beaugrand:2011fd]. Environmental variables utilized in the models included sea surface temperature, surface pH, surface oxygen concentration, and vertically integrated (0–100 m) net primary production (NPP) [@Asch:2018ca]. The ENM-Nereus employed all of the observational data-sources previously mentioned. Global environmental conditions were obtained and results averaged from three Earth System Models developed by the Geophysical Fluid Dynamics Laboratory (GFDL- https://www.gfdl.noaa.gov/earth-system-model/), the Institute Pierre Simon Laplace (IPSL- www.icmc.ipsl.fr/), and the Max Planck Institute for Meteorology (MPI- www.mpimet.mpg.de/en/science/models/). See [@Asch:2018ca; @Reygondeau:2019uh] for model details.

The SDM-SAU model follows a five-steps process based on species-specific life history information, rather than environmental variables [@Close:2006ux; @Palomares:wg]. For each commercial fish species, the model first uses the FAO major fishing areas and countries EEZs to determine a broad distribution. It then uses life history information to delimit its range within the FAO fishing area (e.g. thermal preference, depth limits). The range is delimited even further by expert-review polygons and compared with that of AquaMaps [@Kaschner:2016tl], OBIS and GBIF occurrence data. The model then determines a species habitat preference based on the assumptions that the relative abundance of a species is determined by the number of habitats in a grid cell and the distance of the species to each habitat, as well as the importance of the habitat to the species size. Finally, the species equatorial submergence (e.g. the latitudinal region where a species is not seen in between poles) is estimated for each species. See [@Close:2006ux; @Palomares:wg] for model details.

### *Catch data*

The previous models combine observational data with a series of biotic and abiotic information to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g. ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - i.e., catches  based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in the FAO data available for each country. Catches are also spatially allocated [@Zeller2016]. Thus, we used the Sea Around Us catch reconstruction database from 2005 to 2014 (last ten yeares of data) as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs (average 2005-2014).

@@

We developed a four-criteria method for determining whether or not a species was transboundary. Only species that meet all criteria were considered as transboundary species and the analysis was done only within the EEZs of coastal states. In some cases, criteria also work as indexes to measure the uncertainty in the analysis as they provided continuos index (e.g. from 0 to 1) rather than a discrete, yes or no, index . All of the analysis was done in the statistical software *R version 3.5.2 (2018-12-20)* with the pacages *tidiverse*[@],*data.table*[@],*wesanderson*[@],*janitor*[@],*spdep*[@], *sf*[@],*sp*[@],*rgdal*[@],*tools*[@],*parallel*[@],*rfishbase*[@],*zoo*[@]. All code is available at https://github.com/jepa/FishForVisa.

@@

We estimated the economic contribution of transboundary species for each country using global ex-vessel price [@Tai2017]. The data includes ex-vessel price from multiple sources and a structured interpolation method (e.g. similar countries, species) to fill in data gaps [@Sumaila:2015uc]. The contribution of transboundary species was done by aggregating the catch of each species that happened within the EEZ of the countries sharing that species. For this study, we did not include catch from areas beyond national jurisdiction. Finally, we use the catch data to determine the exploitation category of each species within a EEZ. Although this method has previously used to estimate stock status [@Grainger:1996tk; @Kleisner:2011wn], the categories presented here are intended to be seen as catch trends, and not status of each species. We only assessed species within each EEZ that had at least ten years of data between the first and last reported landings with at least five consecutive data years. Three categories were drawn from the method depending on the year's catch, the maximum catch over time and the minimum catch after the catch peak (**Table XX**). Finally, we reported the predominant category over the last ten years of catch data (2004-2014).
